---
title: "DDA_w_MS1_aqueous"
author: "Lexie"
date: "2023-04-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Post-processing of IDA (DDA) data 

Use the MetaboKit-DDA module to process IDA data files. MS/MS spectra are matched against multiple reference libraries (NIST, MSDIAL, HMDB, LipidBlast), while MS1-level fingerprinting is also supported. This script demonstrates a workflow for preliminary filtering of MetaboKit-DDA output to generate a shortlist of candidate IDs (from both MS/MS and MS1), which will be used to build a project-specific spectral library for subsequent SWATH data processing in the MASLD study.

# 1. Positive mode

```{r}
rm(list = ls()[])
pdata = read.delim("DDA_data/quant_rt_sep_pos.tsv", as.is = T, check.names = F, header = T, na.strings = c("", " ", "NA"))
```

```{r}
parseScores = function(x) {
  nx = length(x)
  ss = rep(NA, nx)
  for(i in 1:nx) {
    xx = strsplit(x[i], ", ")[[1]]
    if(any(xx != "")) ss[i] = max(as.numeric(xx))   
  }
  ss
}
```

## Filter all compounds by signal-to-noise ratio

```{r}
## remove IDs with high signals in blanks
tmp = pdata[,!grepl("RT|score|CE", colnames(pdata))]
tmp[is.na(tmp)] = 0

d = grep("Pos|pos", colnames(tmp))
for(k in d) tmp[,k] = parseScores(as.character(tmp[,k]))

s2n = 5

blank = apply(tmp[,grep("Blank", colnames(tmp))], 1, function(x)median(as.numeric(x)))
liver = apply(tmp[,grep("NASH_Liver", colnames(tmp))], 1, function(x)median(as.numeric(x))) >= (s2n * blank)
plasma = apply(tmp[,grep("NASH_Plasma", colnames(tmp))], 1, function(x)median(as.numeric(x))) >= (s2n * blank)
cm = apply(tmp[,grep("NASH_CM", colnames(tmp))], 1, function(x)median(as.numeric(x))) >= (s2n * blank)

id = liver | plasma | cm
pdata = pdata[id,]
```

## Filter all compounds by detection frequencies

-   Spectra from mouse heart samples (DDA, same LC-MS platform) were also collected for SWATH data extraction

```{r}

pdata = pdata[,-grep("Blank", colnames(pdata))]

Detection_Freq = function(data, matrix){
    x = apply(sapply(X = c("NASH", matrix, "RT"), FUN = grepl, colnames(data)), 1, FUN = all)
    y = apply(sapply(X = c(matrix, "RT"), FUN = grepl, colnames(data)), 1, FUN = all)
    data = data[,x|y]
    
    n = ncol(data)
    freq = apply(data, 1, function(x)(length(which(x != 0))/n))
    return(freq)
}

detected_liver = Detection_Freq(pdata, "Liver")
detected_plasma = Detection_Freq(pdata, "Plasma")
detected_cm = Detection_Freq(pdata, "CM")
detected_mouse = Detection_Freq(pdata, "uL")

d = grep("Pos", colnames(pdata))
pdata = cbind(pdata[,-d], detected_liver, detected_plasma, 
              detected_cm, detected_mouse, pdata[,d])

```

```{r}

MS2 = pdata$`confidence level` == "MSMS"
MS1 = pdata$`confidence level` == "MS1"
id = rep(NA, nrow(pdata))

# MS2
cutoff = 0.2
liver = pdata$detected_liver[MS2] > cutoff
plasma = pdata$detected_plasma[MS2] > cutoff
cm = pdata$detected_cm[MS2] > cutoff
mouse = pdata$detected_mouse[MS2] > cutoff
id[MS2] = liver | plasma | cm | mouse


# MS1
cutoff = 0.5
liver = pdata$detected_liver[MS1] > cutoff
plasma = pdata$detected_plasma[MS1] > cutoff
cm = pdata$detected_cm[MS1] > cutoff
id[MS1] = liver | plasma #| cm

pdata = pdata[id,]
```

## Filter MS2-level compounds

### - Remove compounds with maximum score \< 0.7

score: dot product

```{r}
pdata_ms2 = pdata[pdata$`confidence level` == "MSMS", ]


gg = grep("score_", colnames(pdata_ms2))
tmp = pdata_ms2[,gg]
tmp[is.na(tmp)] = 0
for(k in 1:ncol(tmp)) tmp[,k] = parseScores(as.character(tmp[,k]))

thres = 0.7

liver = apply(tmp[,grep("NASH_Liver", colnames(tmp))], 1, function(x)max(as.numeric(x))) >= thres
plasma = apply(tmp[,grep("NASH_Plasma", colnames(tmp))], 1, function(x)max(as.numeric(x))) >= thres
cm = apply(tmp[,grep("NASH_CM", colnames(tmp))], 1, function(x)max(as.numeric(x))) >= thres
mouse = apply(tmp[,grep("uL", colnames(tmp))], 1, function(x)max(as.numeric(x))) >= thres


id = liver | plasma | cm | mouse
pdata_ms2 = pdata_ms2[id,]
```

## Filter MS1-level putative compounds

### - Remove IDs with RT \> 18.5 min

```{r}
pdata_ms1 = pdata[pdata$`confidence level` == "MS1", ]
pdata_ms1 = pdata_ms1[pdata_ms1$Median < 18.5, ]
```

### - By peak shape (\>= 8) and peak area (median \> 1000)

```{r}
pdata_ms1 = pdata_ms1[pdata_ms1$`peak_shape (median)` >= 8,]

tmp = pdata_ms1[,!grepl("RT|score|CE", colnames(pdata_ms1))]
tmp[is.na(tmp)] = 0

d = grep("Pos", colnames(tmp))
for(k in d) tmp[,k] = parseScores(as.character(tmp[,k]))

liver = apply(tmp[,grep("NASH_Liver", colnames(tmp))], 1, function(x)median(as.numeric(x))) > 1000
plasma = apply(tmp[,grep("NASH_Plasma", colnames(tmp))], 1, function(x)median(as.numeric(x))) > 1000
id = liver | plasma

pdata_ms1 = pdata_ms1[id,]
```

### - Remove isotopes

```{r}

remove_isotope = function(data){
  # identify isotope pairs ( Same RT, mass difference 1.003 Da )
  iso_tol = 0.001
  RT_tol = 0.05
  
  mz_list = data$`feature_m/z` #mz_list isotopic ion
  RT_list = data$Median
  row = rep(NA, nrow(data))
  pair_ind = rep("", nrow(data))
  n = 0
  
  for (i in 1:length(mz_list)){
    mz_diff = mz_list[i] - data$`feature_m/z`
    RT_diff = data$Median - RT_list[i]
    
    s = which(mz_diff < (1.003 + iso_tol) & mz_diff > (1.003 - iso_tol) & abs(RT_diff) < RT_tol)
    
    if (!identical(s, integer(0))){
      n = n + 1  # pair n
      row[i] = i # isotopic ion
      pair_ind[i] = paste0(paste0("isotope_", n), 
                           ifelse(pair_ind[i] != "", paste0(" | ", pair_ind[i]), ""))
      row[s] = s # parent ion
      pair_ind[s] = paste0(paste0("parent_", n), 
                           ifelse(pair_ind[s] != "", paste0(" | ", pair_ind[s]), ""))
    }
  }
  
  df = cbind(row, pair_ind)
  df = na.omit(df) # potential isotopic pairs
  
  
  # Annotate isotopic ratio : carbon * 1.1 % ?
  s = grepl("NASH", colnames(data)) & !grepl("RT_|score_|CE_", colnames(data)) 
  peak = data[,s]
  
  peak[is.na(peak)] = 0 # select a single value of peak area 
  for(k in 1:ncol(peak))peak[,k] = parseScores(as.character(peak[,k])) 
  peak[peak == 0] = NA
  
  ratio_median = isotopes = rep(NA, nrow(peak))
  pair_ind = strsplit(df[,2], " | ")
  
  for (i in 1:n){
    x = which(unlist(lapply(pair_ind, function(x)paste0("parent_", i) %in% x)))
    y = which(unlist(lapply(pair_ind, function(x)paste0("isotope_", i) %in% x)))
    
    p = as.numeric(df[x, 1])
    iso = as.numeric(df[y, 1])
    
    isotopes[p] = df[x, 2]
    isotopes[iso] = df[y, 2]
    
    d = rbind(peak[p,], peak[iso,])
    complete = apply(d, 2, function(x) !any(is.na(x)))
    
    d = data.frame(d[,complete])
    ratio = as.numeric(d[2,]) / as.numeric(d[1,]) / 0.011 
    ratio_median[iso] = median(ratio)
  }
  
  # IDs with very high ratio are less prone to be isotopes
  Non_iso = which(ratio_median > data$`feature_m/z`/12)
  x = grep("isotope", isotopes)
  x = x[!x %in% Non_iso]
  
  data = data[-x, ] # remove isotopes
}

```

```{r}
pdata_ms1 = remove_isotope(pdata_ms1)
```

### - Keep compounds with HMDB/LM accession

```{r}
x = !is.na(pdata_ms1$`accession (HMDB)`) | !is.na(pdata_ms1$`lm_id (LMSD)`)

pdata_ms1 = pdata_ms1[x, ]
```

```{r}
#Combine MS2-level and MS1-level identificaitions

pdata2 = rbind(pdata_ms2, pdata_ms1)
pdata2 = cbind("Add_to_lib" = rep("Y", nrow(pdata2)), pdata2)

write.table(pdata2, "results/dda_postcleaning_pos.txt", sep = "\t", quote = F, row.names = F, na = "")
```

# 2. Negative mode

```{r}
ndata = read.delim("DDA_data/quant_rt_sep_neg.tsv", as.is = T, check.names = F, header = T, na.strings = c("", " ", "NA"))
```

```{r}
x = grep("Phosphatidylcholine lyso 20", ndata$name)
ndata = ndata[-x,]
```

## Filter all compounds by S-to-N ratio

```{r}
tmp = ndata[,!grepl("RT|score|CE", colnames(ndata))]
tmp[is.na(tmp)] = 0

d = grep("Neg|neg", colnames(tmp))
for(k in d) tmp[,k] = parseScores(as.character(tmp[,k]))

s2n = 5

blank = apply(tmp[,grep("Blank", colnames(tmp))], 1, function(x)median(as.numeric(x)))
liver = apply(tmp[,grep("NASH_Liver", colnames(tmp))], 1, function(x)median(as.numeric(x))) >= (s2n * blank)
plasma = apply(tmp[,grep("NASH_Plasma", colnames(tmp))], 1, function(x)median(as.numeric(x))) >= (s2n * blank)
cm = apply(tmp[,grep("NASH_CM", colnames(tmp))], 1, function(x)median(as.numeric(x))) >= (s2n * blank)

id = liver | plasma | cm
ndata = ndata[id,]
```

## Filter all compounds by detection frequencies

```{r}
ndata = ndata[,-grep("Blank", colnames(ndata))]

detected_liver = Detection_Freq(ndata, "Liver")
detected_plasma = Detection_Freq(ndata, "Plasma")
detected_cm = Detection_Freq(ndata, "CM")
detected_mouse = Detection_Freq(ndata, "uL")

d = grep("Neg|neg", colnames(ndata))
ndata = cbind(ndata[,-d], detected_liver, detected_plasma, 
              detected_cm, detected_mouse, ndata[,d])

```

```{r}
MS2 = ndata$`confidence level` == "MSMS"
MS1 = ndata$`confidence level` == "MS1"
id = rep(NA, nrow(ndata))

# MS2
cutoff = 0.2
liver = ndata$detected_liver[MS2] > cutoff
plasma = ndata$detected_plasma[MS2] > cutoff
cm = ndata$detected_cm[MS2] > cutoff
mouse = ndata$detected_mouse[MS2] > cutoff

id[MS2] = liver | plasma | cm | mouse


# MS1
cutoff = 0.5
liver = ndata$detected_liver[MS1] > cutoff
plasma = ndata$detected_plasma[MS1] > cutoff
cm = ndata$detected_cm[MS1] > cutoff

id[MS1] = liver | plasma #| cm

ndata = ndata[id,]
```

## Filter MS2-level compounds

### - Remove compounds with maximum score \< 0.7

```{r}
ndata_ms2 = ndata[ndata$`confidence level` == "MSMS", ]


gg = grep("score_", colnames(ndata_ms2))
tmp = ndata_ms2[,gg]
tmp[is.na(tmp)] = 0
for(k in 1:ncol(tmp)) tmp[,k] = parseScores(as.character(tmp[,k]))

thres = 0.7

liver = apply(tmp[,grep("NASH_Liver", colnames(tmp))], 1, function(x)max(as.numeric(x))) >= thres
plasma = apply(tmp[,grep("NASH_Plasma", colnames(tmp))], 1, function(x)max(as.numeric(x))) >= thres
cm = apply(tmp[,grep("NASH_CM", colnames(tmp))], 1, function(x)max(as.numeric(x))) >= thres
mouse = apply(tmp[,grep("uL", colnames(tmp))], 1, function(x)max(as.numeric(x))) >= thres


id = liver | plasma | cm | mouse
ndata_ms2 = ndata_ms2[id,]
```

## Filter MS1-level putative compounds

### - Remove IDs with RT \> 18.5 min

```{r}
ndata_ms1 = ndata[ndata$`confidence level` == "MS1", ]
ndata_ms1 = ndata[ndata$Median < 18.5, ]
```

### - By peak shape (\>= 8) and peak area (median \> 1000)

```{r}

ndata_ms1 = ndata_ms1[ndata_ms1$`peak_shape (median)` >= 8,]

tmp = ndata_ms1[,!grepl("RT|score|CE", colnames(ndata_ms1))]
tmp[is.na(tmp)] = 0

d = grep("Neg", colnames(tmp))
for(k in d) tmp[,k] = parseScores(as.character(tmp[,k]))

liver = apply(tmp[,grep("NASH_Liver", colnames(tmp))], 1, function(x)median(as.numeric(x))) > 1000
plasma = apply(tmp[,grep("NASH_Plasma", colnames(tmp))], 1, function(x)median(as.numeric(x))) > 1000
id = liver | plasma

ndata_ms1 = ndata_ms1[id,]
```

### - Remove isotopes

```{r}
ndata_ms1 = remove_isotope(ndata_ms1)
```

### - Keep compounds with HMDB/LM accession

```{r}
x = !is.na(ndata_ms1$`accession (HMDB)`) | !is.na(ndata_ms1$`lm_id (LMSD)`)

ndata_ms1 = ndata_ms1[x, ]
```

```{r}

#Combine MS2-level and MS1-level identificaitions

ndata2 = rbind(ndata_ms2, ndata_ms1)
ndata2 = cbind("Add_to_lib" = rep("Y", nrow(ndata2)), ndata2)

write.table(ndata2, "results/dda_postcleaning_neg.txt", sep = "\t", quote = F, row.names = F, na = "")
```

# 
