---
title: "SWATH_data_QC"
author: "Lexie"
date: "2023-04-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Post-processing of SWATH (DIA) data

Filtered MS/MS spectra and MS1-level identifications were used for SWATH data processing with the MetaboKit-DIA module. This script demonstrates a workflow for post-processing the MetaboKit-DIA output.

```{r}
rm(list = ls()[])
source("functions_dia.r")

library(dplyr)
library(readxl)
```

```{r}

pdata = read.delim("SWATH_data/quant_fill_all_pos.tsv", header = T, as.is = T, check.names = F, na = c("", " ", NA))
ndata = read.delim("SWATH_data/quant_fill_all_neg.tsv", header = T, as.is = T, check.names = F, na = c("", " ", NA))

pdata$name = gsub(" \\(E\\)", "", pdata$name)
ndata$name = gsub(" \\(E\\)", "", ndata$name)

names(pdata)[1] = "Var.2"
names(ndata)[1] = "Var.2"
```

Label confidence levels for each entry

```{r}
num = pdata$Var.2
Confidence_level = ifelse(duplicated(num)|duplicated(num, fromLast = T), "MSMS", "MS1")
pdata2 = cbind(Confidence_level, pdata)

num = ndata$Var.2
Confidence_level = ifelse(duplicated(num)|duplicated(num, fromLast = T), "MSMS", "MS1")
ndata2 = cbind(Confidence_level, ndata)
```

# 1. Positive mode

### a) Calculate CoV for QC samples

```{r}
############ liver
Liver_QCarea = pdata2[, QC_area("Liver", pdata2)] 
CoV_liver = CoV_QC(Liver_QCarea)
CoV_liver[CoV_liver == "NaN"] = NA 

############ plasma
Plasma_QCarea = pdata2[, QC_area("Plasma", pdata2)] 
CoV_plasma = CoV_QC(Plasma_QCarea)
CoV_plasma[CoV_plasma == "NaN"] = NA 

############ CM
CM_QCarea = pdata2[, QC_area("CM", pdata2)] 
CoV_cm = CoV_QC(CM_QCarea)
CoV_cm[CoV_cm == "NaN"] = NA 
```

### b) Calculate non-zero rates

Peak area 0: No peak found, area under XIC is zero

```{r}
############ liver
Liver_area = pdata2[,Sample_area("Liver", pdata2)]
n = ncol(Liver_area)
nonzero_rate_liver = apply(Liver_area, 1, function(x)(length(which(x != 0))/n * 100))

############ plasma
Plasma_area = pdata2[,Sample_area("Plasma", pdata2)]
n = ncol(Plasma_area)
nonzero_rate_plasma = apply(Plasma_area, 1, function(x)(length(which(x != 0))/n * 100))

############ CM
CM_area = pdata2[,Sample_area("CM", pdata2)]
n = ncol(CM_area)
nonzero_rate_cm = apply(CM_area, 1, function(x)(length(which(x != 0))/n * 100))
```

### c) Calculate detection frequencies

```{r}
detected_liver = Detection_Freq(pdata2, "Liver")
detected_plasma = Detection_Freq(pdata2, "Plasma")
detected_cm = Detection_Freq(pdata2, "CM")
detected_all = Detection_Freq(pdata2, "")
```

```{r}
pdata3 = cbind(pdata2[,-grep("SWATH", colnames(pdata2))],
               CoV_liver, CoV_plasma, CoV_cm,
               detected_all, detected_liver, 
               detected_plasma, detected_cm,
               nonzero_rate_liver, nonzero_rate_plasma, nonzero_rate_cm,
               pdata2[,grep("SWATH", colnames(pdata2))])
```

### d) Select the best entry as the quantifier

```{r}
# Keep precursor ion and good fragment ions only
pdata3 = select_fragment(pdata3)
# Select the best transition based on CoV
pdata3 = quantifier(pdata3)
```

# 2. Negative mode

### a) Calculate CoV for QC samples

```{r}
############ liver
Liver_QCarea = ndata2[, QC_area("Liver", ndata2)] 
CoV_liver = CoV_QC(Liver_QCarea)
CoV_liver[CoV_liver == "NaN"] = NA 

############ plasma
Plasma_QCarea = ndata2[, QC_area("Plasma", ndata2)] 
CoV_plasma = CoV_QC(Plasma_QCarea)
CoV_plasma[CoV_plasma == "NaN"] = NA 

############ CM
CM_QCarea = ndata2[, QC_area("CM", ndata2)] 
CoV_cm = CoV_QC(CM_QCarea)
CoV_cm[CoV_cm == "NaN"] = NA 
```

### b) Calculate non-zero rates

```{r}
############ liver
Liver_area = ndata2[,Sample_area("Liver", ndata2)]
n = ncol(Liver_area)
nonzero_rate_liver = apply(Liver_area, 1, function(x)(length(which(x != 0))/n * 100))


############ plasma
Plasma_area = ndata2[,Sample_area("Plasma", ndata2)]
n = ncol(Plasma_area)
nonzero_rate_plasma = apply(Plasma_area, 1, function(x)(length(which(x != 0))/n * 100))


############ CM
CM_area = ndata2[,Sample_area("CM", ndata2)]
n = ncol(CM_area)
nonzero_rate_cm = apply(CM_area, 1, function(x)(length(which(x != 0))/n * 100))

```

### c) Calculate detection frequencies

```{r}
detected_liver = Detection_Freq(ndata2, "Liver")
detected_plasma = Detection_Freq(ndata2, "Plasma")
detected_cm = Detection_Freq(ndata2, "CM")
detected_all = Detection_Freq(ndata2, "")
```

```{r}
ndata3 = cbind(ndata2[,-grep("SWATH", colnames(ndata2))],
               CoV_liver, CoV_plasma, CoV_cm,
               detected_all, detected_liver, 
               detected_plasma, detected_cm,
               nonzero_rate_liver, nonzero_rate_plasma, nonzero_rate_cm,
               ndata2[,grep("SWATH", colnames(ndata2))])
```

### d) Select the best entry as the quantifier

```{r}
ndata3 = select_fragment(ndata3)
ndata3 = quantifier(ndata3)
```

# 3. Merge Pos & Neg

```{r}

colnames(pdata3) = gsub("NASH_|A_Pos_SWATH_|A_Pos_Repeat_SWATH_", "", colnames(pdata3))
colnames(pdata3) = gsub("_r2", "", colnames(pdata3))

colnames(ndata3) = gsub("NASH_|A_Neg_SWATH_|A_Neg_Repeat_SWATH_", "", colnames(ndata3))
colnames(ndata3) = gsub("_r2", "", colnames(ndata3))


c = intersect(colnames(pdata3), colnames(ndata3))

pdata3 = pdata3[,c]
ndata3 = ndata3[,c]

Batch = c(rep("HILIC_Pos", nrow(pdata3)), rep("HILIC_Neg", nrow(ndata3)))

met_data = rbind(pdata3, ndata3)
met_data = cbind(Batch, met_data)

x = grep("score_|RT_", colnames(met_data))
met_data = met_data[,-x]

```

### Filter the combined data by detection frequency and CoV

```{r}

tmp = met_data[,grep("CoV", colnames(met_data))]
tmp[is.na(tmp)] = 1e5
  
x = tmp$CoV_liver < 30 | tmp$CoV_plasma < 30 #| tmp$CoV_cm < 30
met_data = met_data[x,]


ms1 = which(met_data$Confidence_level == "MS1")
ms2 = which(met_data$Confidence_level == "MSMS")

cutoff1 = met_data$detected_liver[ms1] < 30 & met_data$detected_plasma[ms1] < 30
cutoff2 = met_data$detected_liver[ms2] < 10 & met_data$detected_plasma[ms2] < 10

rid1 = ms1[cutoff1]
rid2 = ms2[cutoff2]

met_data = met_data[-c(rid1, rid2),]
```

### Name cleaning

```{r}
# match compound name from HMDB or shorter names
met_data = select_name(met_data)

# Lipid names -- keep sum composition nomenclature
met_data$Compound_name = sapply(met_data$Compound_name, function(x)Simplify_lip(x))

# shorter name from HMDB
for (i in 1:nrow(met_data)){
  
  name = met_data$Compound_name[i]
  name_hmdb = met_data[i, "name (HMDB)"]
  
  if (!is.na(name_hmdb) & (nchar(name) > nchar(name_hmdb))){
    met_data$Compound_name[i] = name_hmdb
  } else if (tolower(name) %in% tolower(name_hmdb)) {
    met_data$Compound_name[i] = name_hmdb
  }
}
```

# 4. Mark compounds detected with different adducts (similar RT)

```{r}
met_data2 = met_data %>% group_by(Compound_name) %>% do(add_annot(., 0.33))
```

For the same compound detected with both by MSMS-based IDs and MS1-based IDs:

-   If RTs are different, keep the one with "MSMS" level confidence

```{r}
met_data2 = remove_ms1_diffRT(met_data2)

# repeat the annotation step to identify compounds with different adducts / different RT
met_data2 = within(met_data2, rm("note"))
met_data2 = met_data2 %>% group_by(Compound_name) %>% do(add_annot(., 0.33))
```

```{r}
# Exclude compounds from human exposome
hmdb_exp = read.csv("results/hmdb_exposome_short.csv", as.is = T, check.names = F, header = T)
id = lapply(met_data2$`accession (HMDB)`, function(x)strsplit(x, ","))
a = intersect(unlist(id), hmdb_exp$accession)#intersect of exposome and this dataset

exp = unlist(lapply(id, function(x)all(x %in% a)))
exp = exp & is.na(met_data2$`lm_id (LMSD)`)
met_data2$note[exp] = "human_exposome"

# For manual cleaning
write.table(met_data2, "results/metabolomics_SWATH_for_manual_cleaning.txt", sep = "\t", quote = F, na = "", row.names = F)
```

# 5. After manual cleaning

```{r}
met_data2 = read_excel("results/metabolomics_SWATH_manually_cleaned.xlsx")

rid = c(grep("remove", met_data2$note), grep("human_exposome", met_data2$note))
met_data2 = met_data2[-rid,]

met_data2$name = gsub("(ﾂｬﾂｱ)", "", met_data2$name, fixed = T)
met_data2$name = gsub("(ﾂｬﾂｨﾂｬﾂｱ)", "", met_data2$name, fixed = T)
met_data2$`name (HMDB)` = gsub("(ﾂｬﾂｱ)", "", met_data2$`name (HMDB)`, fixed = T)
met_data2$`name (HMDB)` = gsub("(ﾂｬﾂｨﾂｬﾂｱ)", "", met_data2$`name (HMDB)`, fixed = T)

met_data2 = within(met_data2, rm("note"))
met_data2 = met_data2 %>% group_by(Compound_name) %>% do(add_annot(., 0.33))
table(met_data2$note)
```

For the same compound with multiple adducts:

-   For each tissue, sum up (geometric mean) intensity if the entry has CoV \< 30 & detection frequency \> 30

```{r}
x = c("Var.2", "MS_level", "precursor_m/z", "fragment_m/z")

met_data3 = met_data2 %>% 
  mutate(across(all_of(x), ~as.character(.))) %>% 
  group_by(Compound_name, note) %>% 
  do(
    if (all(grepl("DiffAdduct", .$note))) {
      concat_adduct(.)
    } else {
      .
      }
  )
```

-   Re-calculate QC CoV for compounds with intensity summing up

```{r}
# Liver
QCarea = met_data3[, grep("Liver_QC", colnames(met_data3))]
x = which(met_data3$CoV_liver == 1e7)
all(is.na(CoV_QC(QCarea)[x]) |CoV_QC(QCarea)[x] < 30 ) # TRUE
met_data3$CoV_liver[x] = CoV_QC(QCarea)[x]

# Plasma
QCarea = met_data3[, grep("Plasma_QC", colnames(met_data3))]
x = which(met_data3$CoV_plasma == 1e7)
all(is.na(CoV_QC(QCarea)[x]) |CoV_QC(QCarea)[x] < 30 ) # TRUE
met_data3$CoV_plasma[x] = CoV_QC(QCarea)[x]

# CM
QCarea = met_data3[, grep("CM_QC", colnames(met_data3))]
x = which(met_data3$CoV_cm == 1e7)
all(is.na(CoV_QC(QCarea)[x]) |CoV_QC(QCarea)[x] < 30 ) # TRUE
met_data3$CoV_cm[x] = CoV_QC(QCarea)[x]
```

-   Label compounds with multiple RTs

```{r}
met_data3$note = NA
mul_RT = (duplicated(met_data3$Compound_name) | duplicated(met_data3$Compound_name, fromLast = T))
met_data3$note = ifelse(mul_RT, "Multi_RT", met_data3$note)

write.table(met_data3, "results/metabolomics_SWATH_data.txt", sep = "\t", quote = F, na = "", row.names = F)
```

Repeat the same workflow for lipidomics data (organic fraction).

Combine the two parts and perform downstream analysis.
