---
title: "SWATH_data_QC"
author: "Lexie"
date: "2023-04-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Post-processing of SWATH (DIA) data 

Filtered MS/MS spectra and MS1-level identifications were used for SWATH data processing with the MetaboKit DIA module. This script demonstrates a workflow for post-processing the MetaboKit-DIA output.

```{r}

# Swath
pdata = read.delim("SWATH_data/quant_fill_all_pos.tsv", header = T, as.is = T, check.names = F, na = c("", " ", NA))
ndata = read.delim("SWATH_data/quant_fill_all_neg.tsv", header = T, as.is = T, check.names = F, na = c("", " ", NA))

pdata$name = gsub(" \\(E\\)", "", pdata$name)
ndata$name = gsub(" \\(E\\)", "", ndata$name)

names(pdata)[1] = "Var.2"
names(ndata)[1] = "Var.2"
```

Label confidence levels for each entry

```{r}
num = pdata$Var.2
Confidence_level = ifelse(duplicated(num)|duplicated(num, fromLast = T), "MSMS", "MS1")
pdata2 = cbind(Confidence_level, pdata)

num = ndata$Var.2
Confidence_level = ifelse(duplicated(num)|duplicated(num, fromLast = T), "MSMS", "MS1")
ndata2 = cbind(Confidence_level, ndata)
```

# 1. Positive mode

### a) Calculate CoV for QC samples

```{r}
################
#### Functions
################

# Get QC quant data for the matrix
QC_area = function(matrix, data){
  x = apply(sapply(X = c("NASH", "_QC", matrix), FUN = grepl, colnames(data)), 1, FUN = all)
  y = !grepl("RT_|score_", colnames(data)) 
  return(x & y)
}

# normalize QC 
normalizeQC = function(x){
  sum = apply(x, 2, function(x) sum(x, na.rm=TRUE))
  x2 = x
  for(j in 1:ncol(x2)) {
    x2[,j] = x2[,j]/sum[j]
  }
  x2
}

# Calculate CoV
CoV_QC = function(data){
  nor_d = normalizeQC(data)
  CoV = rep(NA, nrow(data))
  for(i in 1:nrow(data)){
    x = as.numeric(nor_d[i,])
    xx = x[!is.na(x)]
    if(length(xx) >= 3) {
      CoV[i] = sd(xx)/mean(xx)*100
    }
  }
  return(CoV)
}
```

```{r}
############ liver
Liver_QCarea = pdata2[, QC_area("Liver", pdata2)] 
CoV_liver = CoV_QC(Liver_QCarea)
CoV_liver[CoV_liver == "NaN"] = NA 


############ plasma
Plasma_QCarea = pdata2[, QC_area("Plasma", pdata2)] 
CoV_plasma = CoV_QC(Plasma_QCarea)
CoV_plasma[CoV_plasma == "NaN"] = NA 


############ CM
CM_QCarea = pdata2[, QC_area("CM", pdata2)] 
CoV_cm = CoV_QC(CM_QCarea)
CoV_cm[CoV_cm == "NaN"] = NA 
```

### b) Calculate non-zero rates

```{r}

Sample_area = function(matrix, data){
  x = apply(sapply(X = c("NASH",  matrix), FUN = grepl, colnames(data)), 1, FUN = all)
  y = !grepl("RT_|score_|_QC", colnames(data)) 
  return(x & y)
}

############ liver
Liver_area = pdata2[,Sample_area("Liver", pdata2)]
n = ncol(Liver_area)
nonzero_rate_liver = apply(Liver_area, 1, function(x)(length(which(x != 0))/n * 100))


############ plasma
Plasma_area = pdata2[,Sample_area("Plasma", pdata2)]
n = ncol(Plasma_area)
nonzero_rate_plasma = apply(Plasma_area, 1, function(x)(length(which(x != 0))/n * 100))


############ CM
CM_area = pdata2[,Sample_area("CM", pdata2)]
n = ncol(CM_area)
nonzero_rate_cm = apply(CM_area, 1, function(x)(length(which(x != 0))/n * 100))


```

### c) Calculate detection frequencies

```{r}

Detection_Freq = function(data, matrix){
    x = apply(sapply(X = c("NASH", matrix, "RT"), FUN = grepl, colnames(data)), 1, FUN = all)
    y = !grepl("QC_", colnames(data))
    data = data[,x & y]
    
    n = ncol(data)
    freq = apply(data, 1, function(x)(length(which(x != 0))/n * 100))
    return(freq)
}

detected_liver = Detection_Freq(pdata2, "Liver")
detected_plasma = Detection_Freq(pdata2, "Plasma")
detected_cm = Detection_Freq(pdata2, "CM")
detected_all = Detection_Freq(pdata2, "")

```

```{r}
pdata3 = cbind(pdata2[,-grep("SWATH", colnames(pdata2))],
               CoV_liver, CoV_plasma, CoV_cm,
               detected_all, detected_liver, 
               detected_plasma, detected_cm,
               nonzero_rate_liver, nonzero_rate_plasma, nonzero_rate_cm,
               pdata2[,grep("SWATH", colnames(pdata2))])
```

### d) Select the best transition as the quantifier

Select precursor ion and good fragment ions only

```{r}

select_fragment = function(data){
  ind = unique(data$Var.2)
  nr = 1:nrow(data)
  nr_new = NULL
  
  # 1) remove the fragment ions with frequencies < 0.9 of the precursor ion's freq
  for (i in 1:length(ind)){
    cpd = which(data$Var.2 == ind[i])
    d = data[cpd,]
    ms1 = which(d$MS_level == "1")
    ms2 = which(d$MS_level == "2")
    x = which(d$detected_all[ms2] >= d$detected_all[ms1] * 0.9)
    nr_new = append(nr_new, c(cpd[ms1], cpd[ms2][x]))
    }
  data = data[nr_new,]
  
  # 2) filter fragment ions by criteria "correlation > 0.8" 
  tmp = data[,grep("score_", colnames(data))]
  Corr_median = apply(tmp, 1, function(x)median(x, na.rm = T))
  f = which(data$MS_level == "2")
  Corr_median[f] = sapply(Corr_median[f], function(x)ifelse(x > 0.8, x, NA))
  
  data = data[!is.na(Corr_median),]
  return(data)
}

pdata3 = select_fragment(pdata3)
```

Select the best transition based on CoV

```{r}

quantifier = function(data){
  ind = unique(data$Var.2)
  nr = 1:nrow(data)
  nr_new = NULL
  
  for (i in 1:length(ind)){
    cpd = which(data$Var.2 == ind[i])
    d = data[cpd,]
    
    freq = grep("detected_liver|detected_plasma|detected_cm", colnames(d))
    
    # If there are equal values, 
    # "which.max" will take the first one (priority: liver > plasma > cm)
    m = strsplit(names(which.max(d[1, freq])), "_")[[1]][2]
    cov = apply(sapply(X = c("CoV",  m), FUN = grepl, colnames(d)), 1, all)
    nr_new = append(nr_new, cpd[which.min(d[,cov])])
  }
  
  data = data[nr_new,]
  return(data)
}

pdata3 = quantifier(pdata3)
```

# 2. Negative mode

### a) Calculate CoV for QC samples

```{r}
############ liver
Liver_QCarea = ndata2[, QC_area("Liver", ndata2)] 
CoV_liver = CoV_QC(Liver_QCarea)
CoV_liver[CoV_liver == "NaN"] = NA 


############ plasma
Plasma_QCarea = ndata2[, QC_area("Plasma", ndata2)] 
CoV_plasma = CoV_QC(Plasma_QCarea)
CoV_plasma[CoV_plasma == "NaN"] = NA 


############ CM
CM_QCarea = ndata2[, QC_area("CM", ndata2)] 
CoV_cm = CoV_QC(CM_QCarea)
CoV_cm[CoV_cm == "NaN"] = NA 
```

### b) Calculate non-zero rates

```{r}
############ liver
Liver_area = ndata2[,Sample_area("Liver", ndata2)]
n = ncol(Liver_area)
nonzero_rate_liver = apply(Liver_area, 1, function(x)(length(which(x != 0))/n * 100))


############ plasma
Plasma_area = ndata2[,Sample_area("Plasma", ndata2)]
n = ncol(Plasma_area)
nonzero_rate_plasma = apply(Plasma_area, 1, function(x)(length(which(x != 0))/n * 100))


############ CM
CM_area = ndata2[,Sample_area("CM", ndata2)]
n = ncol(CM_area)
nonzero_rate_cm = apply(CM_area, 1, function(x)(length(which(x != 0))/n * 100))

```

### c) Calculate detection frequencies

```{r}
detected_liver = Detection_Freq(ndata2, "Liver")
detected_plasma = Detection_Freq(ndata2, "Plasma")
detected_cm = Detection_Freq(ndata2, "CM")
detected_all = Detection_Freq(ndata2, "")
```

```{r}
ndata3 = cbind(ndata2[,-grep("SWATH", colnames(ndata2))],
               CoV_liver, CoV_plasma, CoV_cm,
               detected_all, detected_liver, 
               detected_plasma, detected_cm,
               nonzero_rate_liver, nonzero_rate_plasma, nonzero_rate_cm,
               ndata2[,grep("SWATH", colnames(ndata2))])
```

### d) Select the best transition as the quantifier

```{r}
ndata3 = select_fragment(ndata3)
ndata3 = quantifier(ndata3)
```

# 3. Merge Pos & Neg

```{r}

colnames(pdata3) = gsub("NASH_|A_Pos_SWATH_|A_Pos_Repeat_SWATH_", "", colnames(pdata3))
colnames(pdata3) = gsub("_r2", "", colnames(pdata3))

colnames(ndata3) = gsub("NASH_|A_Neg_SWATH_|A_Neg_Repeat_SWATH_", "", colnames(ndata3))
colnames(ndata3) = gsub("_r2", "", colnames(ndata3))


c = intersect(colnames(pdata3), colnames(ndata3))

pdata3 = pdata3[,c]
ndata3 = ndata3[,c]

Batch = c(rep("HILIC_Pos", nrow(pdata3)), rep("HILIC_Neg", nrow(ndata3)))

met_data = rbind(pdata3, ndata3)
met_data = cbind(Batch, met_data)

x = grep("score_|RT_", colnames(met_data))
met_data = met_data[,-x]

```

### Filter the combined data by detection frequency and CoV

```{r}

tmp = met_data[,grep("CoV", colnames(met_data))]
tmp[is.na(tmp)] = 1e5
  
x = tmp$CoV_liver < 30 | tmp$CoV_plasma < 30 #| tmp$CoV_cm < 30
met_data = met_data[x,]


ms1 = which(met_data$Confidence_level == "MS1")
ms2 = which(met_data$Confidence_level == "MSMS")

cutoff1 = met_data$detected_liver[ms1] < 30 & met_data$detected_plasma[ms1] < 30
cutoff2 = met_data$detected_liver[ms2] < 10 & met_data$detected_plasma[ms2] < 10

rid1 = ms1[cutoff1]
rid2 = ms2[cutoff2]

met_data = met_data[-c(rid1, rid2),]
```

### Name cleaning

```{r}
name_split = function(string, adduct){
  Adduct = unique(unlist(strsplit(names(table(adduct)), ", ")))
  ion = paste0(" ", paste0(Adduct, collapse = "| "))
  ion = gsub("+", "_", ion, fixed =T)
  n = strsplit(as.character(string), " --- |possibly an ISF --- ")    
  nn = unlist(n)
  
  for (i in 1:length(nn)) {  
    nn[i] = gsub("+", "_", nn[i], fixed =T)
    nn[i] = strsplit(nn[i], ion)[[1]][1]
    nn[i] = strsplit(nn[i], "; | cation| Cation")[[1]][1]  # "; " for annotations in MassBank
  }
  string_new = paste0(na.omit(unique(nn)), collapse = " --- ")
  return(string_new)
}

select_name = function(data){
  library(tibble)
  library(dplyr)
  library(stringr)
  
  name = sapply(data$name, function(x)name_split(x, data$adduct))
  f = rep(NA, nrow(data))
  
  for (i in 1:nrow(data)){
    name_in_lib = as.character(name[i])
    name_hmdb = data[i, "name (HMDB)"]
    abbrev = data[i, "abbrev (LMSD)"]
    if(!is.na(abbrev)){
      f[i] = abbrev
    } #else if (tolower(name_in_lib) %in% tolower(name_hmdb)) {
      #f[i] = name_hmdb
    #} #else if (!is.na(name_hmdb) & (nchar(name_in_lib) > nchar(name_hmdb))){
      #f[i] = name_hmdb
    #} 
    else {
      f[i] = name_in_lib
    }
  }
  
  data = add_column(data, Compound_name = f, .before = "name") %>% arrange(Compound_name)
  return(data)
}
```

```{r}
met_data = select_name(met_data)
```

Lipid names -- keep sum composition nomenclature

```{r}
library(stringr)
library(stringi)
library(gsubfn)

Lnames_convert = function(string){
  x = gsub("\\(.*?\\)", "", string) 
  x = gsub("-SN\\d+", "", x)
  matched_colon = gregexpr(":", x, fixed = TRUE)
  n_colon = length(unlist(matched_colon))
  
  if (grepl("ST |FAHFA ", string)){
    x_new = x
    } else if ((n_colon == "1") & !(str_detect(x, "([0-9]+)O"))) {
      x_new = x
      } else{
        x2 = str_split(x, " ")
        species = x2[[1]][1]
        unit = str_split(x2[[1]][2], "_|;|O-|N-|/")
        
        ## remove unit without colon
        unit = lapply(unit, function(x)x[grep(":",x)])
        
        Cdb = str_extract_all(unlist(unit), "[^:]+")
        carbon = sum(as.numeric(lapply(Cdb, function(x)x[1])))
        db = sum(as.numeric(lapply(Cdb, function(x)x[2])))
        O = ifelse(str_detect(x, "O-"), "O-", "")
        N = ifelse(str_detect(x, "N-"), "N-", "")
        
        tmp = gsub(";O", ";1O", x2)
        #convert ";2O" into ";O2"
        if (str_detect(tmp, "([0-9]+)O")){
          tmp = unlist(str_extract_all(tmp, "([0-9]+)O"))
          n = sum(as.integer(gsub("O", "", tmp)))
          n = ifelse(n == 1, "", n)
          nO = paste0(";O", n)
          if (str_detect(string, "OH\\)")){
            nO = gsubfn("\\d+", function(x) as.numeric(x) + 1, nO)
          } 
        } else nO = ""
        x_new = paste0(species," ", O, N, carbon,":",db, nO)
      }
  
  # Add (FA xx) or (O-xx) chain
  if(str_detect(string, "\\(FA|\\(O-")){
    y = unlist(strsplit(string, "\\(|\\)"))
    y = y[grep("FA |O-", y)]
    
    if(length(y) > 0){
      y2 = strsplit(y, " |O-")[[1]][2]
      carbon2 = as.numeric(strsplit(y2, ":")[[1]][1])
      db2 = as.numeric(strsplit(y2, ":")[[1]][2])
      carbon = carbon + carbon2
      db = db + db2
      x_new = paste0(species," ", O, N, carbon,":",db, nO)
    }
    }
  return(x_new)
}

Simplify_lip = function(string){
  n = str_split(string, " --- ")    
  nn = unlist(n)
  
  for (i in 1:length(nn)){
    if (grepl(":", nn[i])){
      nn[i] = Lnames_convert(nn[i])
    }
  }
  nn = unique(nn)
  string_new = paste0(na.omit(nn), collapse = " --- ")
  return(string_new)
}
```

```{r}
met_data$Compound_name = sapply(met_data$Compound_name, function(x)Simplify_lip(x))

# shorter name from HMDB
for (i in 1:nrow(met_data)){
  
  name = met_data$Compound_name[i]
  name_hmdb = met_data[i, "name (HMDB)"]
  
  if (!is.na(name_hmdb) & (nchar(name) > nchar(name_hmdb))){
    met_data$Compound_name[i] = name_hmdb
  } else if (tolower(name) %in% tolower(name_hmdb)) {
    met_data$Compound_name[i] = name_hmdb
  }
}
```

# 4. Mark compounds detected with different adducts (similar RT)

```{r}

add_annot = function(data, tolerance){
  library(data.table)
  library(tibble)
  library(dplyr)
  
  a = data$RT
  a_ = sort(a, index.return = T)
  diff = abs(shift(a_$x) - a_$x)
  
  group = list()
  n = 1
  group[[n]] = a_$ix[1]
  
  if(length(diff) > 1){
    for (i in 2:length(diff)){
      if (diff[i] < tolerance){
        group[[n]] = c(a_$ix[i], unlist(group[n]))
      } else {
        n = n+1
        group[[n]] = a_$ix[i]
      }
    }
  }
  
  library(tibble)
  data = add_column(data, note = rep(NA, nrow(data)), .before = "name")
  x = unlist(lapply(group, function(x)length(x) > 1))
  
  if(any(x)){
    num = 0
    for (i in 1:length(group)){
      ind = unlist(group[[i]])
      if (length(ind) == 1){
        data$note[ind] = "remove"
        } else if (length(ind) > 1 & (sum(x) < 2)){
          data$note[ind] = "DiffAdduct"
          } else if (length(ind) > 1 & (sum(x) >= 2)){
            num = num + 1
            data$note[ind] = paste0("DiffAdduct", num)
          }
      }
    }
  mul_RT = is.na(data$note) &
    (duplicated(data$Compound_name) | duplicated(data$Compound_name, fromLast = T))
  data$note = ifelse(mul_RT, "Multi_RT", data$note)
  return(data)
}
```

```{r}
met_data2 = met_data %>% group_by(Compound_name) %>% do(add_annot(., 0.33))

```

For the same compound detected with both by MSMS-based IDs and MS1-based IDs:

-   If RT are different, keep the one with "MSMS" level confidence

```{r}
remove_ms1_diffRT = function(data){
  id = unique(data$Compound_name)
  rid = NULL
  
  for (i in 1:length(id)){
  ind = which(data$Compound_name == id[i])
  if (length(ind > 1)){
    conf = data[ind, "Confidence_level"]
    multRT = data[ind, "note"]
    if (grepl("MSMS", conf) & grepl("MS1", conf) & grepl("Multi_RT", multRT)){
      rid = append(rid, ind[which(conf == "MS1")])
      }
    }
  }
  data = data[-rid,]
  return(data)
}
```

```{r}
met_data2 = remove_ms1_diffRT(met_data2)

# repeat the annotation step to identify compounds with different adducts / different RT
met_data2 = within(met_data2, rm("note"))
met_data2 = met_data2 %>% group_by(Compound_name) %>% do(add_annot(., 0.33))
```

```{r}
# Exclude compounds from human exposome
hmdb_exp = read.csv("results/hmdb_exposome_short.csv", as.is = T, check.names = F, header = T)

id = lapply(met_data2$`accession (HMDB)`, function(x)strsplit(x, ","))
a = intersect(unlist(id), hmdb_exp$accession)#intersect of exposome and this dataset

exp = unlist(lapply(id, function(x)all(x %in% a)))
exp = exp & is.na(met_data2$`lm_id (LMSD)`)

met_data2$note[exp] = "human_exposome"
```

Manual cleaning

```{r}
write.table(met_data2, "results/metabolomics_SWATH_for_manual_cleaning.txt", sep = "\t", quote = F, na = "", row.names = F)
```

# 5. Post-manual cleaning

Remove compounds

```{r}

library(readxl)
met_data2 = read_excel("results/metabolomics_SWATH_manually_cleaned.xlsx")

rid = c(grep("remove", met_data2$note), grep("human_exposome", met_data2$note))
met_data2 = met_data2[-rid,]

met_data2$name = gsub("(¬±)", "", met_data2$name, fixed = T)
met_data2$name = gsub("(¬¨¬±)", "", met_data2$name, fixed = T)
met_data2$`name (HMDB)` = gsub("(¬±)", "", met_data2$`name (HMDB)`, fixed = T)
met_data2$`name (HMDB)` = gsub("(¬¨¬±)", "", met_data2$`name (HMDB)`, fixed = T)

met_data2 = within(met_data2, rm("note"))
met_data2 = met_data2 %>% group_by(Compound_name) %>% do(add_annot(., 0.33))
table(met_data2$note)
```

For the same compound with multiple adducts:

-   For each tissue, sum up (hypergeometric mean) intensity if the entry has CoV \< 30 & detection frequency \> 30

```{r}

concat_adduct = function(data){
  library(stringr)
  
  concat = data.frame(matrix(NA, nrow = 1, ncol = ncol(data)))
  colnames(concat) = colnames(data)
  
  # Different operation for columns
  # 1. CoV"  1e7 for repeating CoV calculation
  col1 = which(colnames(data) %in% c("CoV_liver", "CoV_plasma", "CoV_cm"))
  # 2. detected_x/nonzero_rate, max
  col2 = c(grep("detected_", colnames(data)), grep("nonzero", colnames(data)))
  # 3. Samples: Sum up
  col3_l = which(grepl("Liver_", colnames(data)) & (str_detect(colnames(data),"[0-9]{3}") | grepl("QC", colnames(data))))
  col3_p = which(grepl("Plasma_", colnames(data)) & (str_detect(colnames(data),"[0-9]{3}") | grepl("QC", colnames(data))))
  col3_c = which(grepl("CM_", colnames(data)) & (str_detect(colnames(data),"[0-9]{3}") | grepl("QC", colnames(data))))
  # 4. RT
  col4 = grep("RT", colnames(data))
  # 5. information: paste with collapse", ", keep unique ones
  col5 = which(! 1:ncol(data) %in% c(col1, col2, col3_l, col3_p, col3_c, col4))
   
  concat[1, col1] = 1e7
  concat[1, col2] = apply(data[, col2, drop = FALSE], 2, max, na.rm = TRUE)
  concat[1, col4] = apply(data[, col4, drop = FALSE], 2, median, na.rm = TRUE)
  concat[1, col5] = apply(data[, col5, drop = FALSE], 2, function(x)ifelse(all(is.na(x)), NA, paste(unique(x[!is.na(x)]), collapse = ", ")))
  concat[1, "note"] = NA  # remove annotation
  
  r = which(data$CoV_liver < 30 & data$detected_liver > 30)
  concat[1, "CoV_liver"]  = ifelse(length(r) == 1, data$CoV_liver[r], 1e7) #CoV"  1e7 for repeating CoV calculation
  concat[1, col3_l] = apply(data[r, col3_l], 2, function(x)ifelse(length(r) >= 1, exp(mean(log(x))), NA))
  
  r = which(data$CoV_plasma < 30 & data$detected_plasma > 30)
  concat[1, "CoV_plasma"]  = ifelse(length(r) == 1, data$CoV_plasma[r], 1e7) 
  concat[1, col3_p] = apply(data[r, col3_p], 2, function(x)ifelse(length(r) >= 1, exp(mean(log(x))), NA))
  
  r = which(data$CoV_cm < 30 & data$detected_cm > 30)
  concat[1, "CoV_cm"]  = ifelse(length(r) == 1, data$CoV_cm[r], 1e7) 
  concat[1, col3_c] = apply(data[r, col3_c], 2, function(x)ifelse(length(r) >= 1, exp(mean(log(x))), NA))
  
  return(concat)
}
```

```{r}

library(dplyr)
x = c("Var.2", "MS_level", "precursor_m/z", "fragment_m/z")

met_data3 = met_data2 %>% mutate(across(all_of(x), ~as.character(.))) %>% 
  group_by(Compound_name, note) %>% 
  do(
    if (all(grepl("DiffAdduct", .$note))) {
      concat_adduct(.)
    } else {
      .
      }
  )
```

Re-calculate QC CoV for compounds with intensity summing up

```{r}
# Liver
QCarea = met_data3[, grep("Liver_QC", colnames(met_data3))]
x = which(met_data3$CoV_liver == 1e7)
all(is.na(CoV_QC(QCarea)[x]) |CoV_QC(QCarea)[x] < 30 ) # TRUE
met_data3$CoV_liver[x] = CoV_QC(QCarea)[x]

# Plasma
QCarea = met_data3[, grep("Plasma_QC", colnames(met_data3))]
x = which(met_data3$CoV_plasma == 1e7)
all(is.na(CoV_QC(QCarea)[x]) |CoV_QC(QCarea)[x] < 30 ) # TRUE
met_data3$CoV_plasma[x] = CoV_QC(QCarea)[x]

# CM
QCarea = met_data3[, grep("CM_QC", colnames(met_data3))]
x = which(met_data3$CoV_cm == 1e7)
all(is.na(CoV_QC(QCarea)[x]) |CoV_QC(QCarea)[x] < 30 ) # TRUE
met_data3$CoV_cm[x] = CoV_QC(QCarea)[x]
```

Label compounds with multiple RTs

```{r}

met_data3$note = NA
mul_RT = (duplicated(met_data3$Compound_name) | duplicated(met_data3$Compound_name, fromLast = T))
met_data3$note = ifelse(mul_RT, "Multi_RT", met_data3$note)

write.table(met_data3, "results/metabolomics_SWATH_data.txt", sep = "\t", quote = F, na = "", row.names = F)
```

Repeat the same workflow for lipidomics data (organic fraction).

Combine the two parts and perform downstream analysis.
